---
csl: assets/apa.csl
bibliography: assets/refs.bib
...

```{r methods-setup, include = FALSE}
library(knitr)

# working dir for evaluating code (use project directory)
wd <- rprojroot::find_rstudio_root_file()
opts_knit$set(root.dir = wd)

opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r methods-data}
# load yaml bits here
library(yaml)
n_counts <- yaml.load_file("./data/screening_facts.yaml")
n_exc_lenas <- n_counts$inadequate_lenas + n_counts$refused_lena
```

Methods and Measurements
------------------------------------------------------------------------

### Participants

We report all measurements and data exclusions following guidelines in
@OSF_Statement. We examined data from the first two time points of a
longitudinal study of preschoolers from English-speaking households. At
Time 1 in the study, the children were 28--39 months old. During Time 1,
we collected our measures of language input, lexical processing and
vocabulary. At Time 2, we collected follow-up vocabulary measures when
the children were 39--52 months old.

A total of `r n_counts$starting` children provided vocabulary,
processing and input data at Time 1. We excluded `r n_counts$cis`
children with cochlear implants from the present analysis. We also
excluded `r n_counts$lts` children identified by parents as
late-talkers. Of the remaining children, `r n_counts$returned_to_t2`
provided vocabulary measures at Time 2. As detailed below, we excluded
`r n_counts$inadequate_lenas + n_counts$refused_lena` children for
having inadequate home-language recordings and
`r n_counts$unreliable_et` children for having unreliable eyetracking
data. A final total of `r n_counts$with_reliable_et` children were used
in the vocabulary analyses. A small subset of the Time 1 vocabulary and
eyetracking data (*n* = `r n_counts$in_other_paper`) was previously
reported in @RWLPaper, which analyzed vocabulary size and concurrent
lexical processing in a diverse group of participants. All children
underwent a hearing screening at both time points, and they had normal
speech, language, and cognitive development according to parent report.

Stimuli were presented in children's home dialect, either Mainstream
American English (MAE) or African American English (AAE). We made an
initial guess about what the home dialect was likely to be based on a
number of factors, including the recruitment source and the child's
address. For most children, the home dialect was MAE. If we thought the
home dialect might be AAE, a native AAE speaker who was a fluent
dialect-shifter was scheduled for the lab visit, and she confirmed the
home dialect by listening to the caregiver interact with the child
during the consent procedure at the beginning of the visit. AAE was the
home dialect for `r n_counts$n_aae_with_reliable_et` of the
`r n_counts$with_reliable_et` participants.

Several other measurements were collected as part of the longitudinal
study. They are not analyzed here because we limit attention to only the
measures relevant for the analysis of input, processing, and vocabulary.
Other unanalyzed Time 1 tasks were two picture-prompted word-repetition
tasks [@RealWordRep2008], an eyetracking task with mispronunciations
of familiar words [@MPPaper], a minimal pair discrimination task
[based on @ProtoMinPair], a verbal fluency task [WJ-III Retrieval
Fluency subtest, @WCJ3], a shape stroop task [@Carlson2005], and an
articulation test [GFTA-2, @GFTA2]. Parents completed the
*MacArthur-Bates Communicative Development Inventory* [@MBCDI], an
inventory about executive function [BRIEF-P, @BRIEFP], a survey about
early literacy practices in the home [@Senechal2006], and a
demographic survey that included a multiple-choice question on maternal
education level. A similar test battery was used at Time 2, with the
addition of new tasks targeting phonological awareness [CTOPP-2,
@CTOPP2] and speech perception [SAILS task in @Rvachew2006].


### Vocabulary

At both time points, children received the *Expressive Vocabulary
Test, 2^nd^ Edition* [EVT-2, @EVT2] and its receptive counterpart, the
*Peabody Picture Vocabulary Test, 4^th^ Edition* [PPVT-4, @PPVT4]. In
the expressive test, children were presented an image and had to name
it. In the receptive test, children were presented four images and had
to select a named image. For our analyses, we used growth scale values
provided by each test; these values transform raw scores (words correct)
into a scale that grows linearly with age.


### Language input

Language input data was collected using a Language Environment Analysis
(LENA) digital recorder, a small device worn by a child [@LENAAudio].
The device records all audible sounds for up to 16 hours. The recorder
and instructions for using it were given to families. We instructed
families to activate the recorder in the morning and record a typical
day for the child. LENA software analyzed each recording to generate a
summary of the child's language environment [@LENAAudio]. The measures
included 1) hourly word-counts of adult language in the child's
environment, 2) hourly number of child-adult and adult-child
conversational turns, 3) hourly proportions of meaningful (nearby)
speech, distant speech, noise, television/electronics, and silence,
and 4) hourly number of child vocalizations.

We computed the averages of each of these hourly measurements, excluding
data from hours recorded after midnight. We computed the duration of the
remaining before-midnight data in seconds, computing the number of hours
from the number of seconds. This adjustment corrects for hours where the
recording started midway through an hour. The average hourly adult word
count then was the total adult word count in the recording divided by
the number of hours. Our procedure differed from that of @Weisleder2013:
That study only used the adult word counts from segments that coders had
classified as child-directed.

We excluded recordings that might provide unreliable information. We
excluded `r n_counts$inadequate_lenas` recordings with less than 10
hours of data recorded before midnight, because such recordings
undersampled the child's day. LENA software documentation also
recommends that recordings be at least 10 hours in duration
[@LENAResearchFoundation2015]. We excluded `r n_counts$refused_lena`
recording from a child who did not wear the device.
