---
csl: assets/apa.csl
bibliography: assets/refs.bib
...

```{r discussion-setup, include = FALSE}
library(knitr)
wd <- rprojroot::find_rstudio_root_file()
opts_knit$set(root.dir = wd)
opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r discussion-data}
# maternal education counts
library(dplyr)
library(yaml)
medu_counts <- yaml.load_file("./data/screening_facts.yaml") %>% 
  getElement("medu_counts_with_reliable_et")

medu_counts$total <- sum(unlist(medu_counts))
d_all <- readr::read_csv("./data/03_input_vocab_eyetracking.csv")
r_pred <- cor(d_all$ot1, d_all$AWC_Hourly) %>% 
  printy::fmt_fix_digits(2) %>% 
  printy::fmt_leading_zero()
```

Discussion
-----------------------------------------------------------------------

We asked how lexical processing efficiency and home language input
predicted vocabulary size one year later in a large sample of
preschoolers. We measured lexical processing using an eyetracking
experiment and model-derived estimates of how quickly on average a
child's gaze shifted to a named image. We measured language input using
the average number of adult words per hour from LENA recordings, and we
measured expressive and receptive vocabulary directly using standardized
tests. We first tested how language input and lexical processing at
age 3 predicted vocabulary size at age 4 without controlling for age-3
vocabulary levels. In these *baseline* analyses, both measures reliably
and independently predicted vocabulary size. The processing effect
was 2–2.5 times larger in magnitude than the input effect. Lexical
processing and language input were weakly correlated, *r* = `r r_pred`,
and there were no credible interaction Processing × Input effects. These
baseline analyses support the conclusions that lexical processing
efficiency was a more important predictor of future vocabulary than
language input and that word recognition efficiency did not constrain
the beneficial effects of language exposure on future vocabulary size.

We next examined how input and lexical processing related to vocabulary
*growth*, by controlling for age-3 vocabulary. The best predictors of
Time 2 expressive vocabulary were Time 1 expressive vocabulary followed
by lexical processing. The processing effect was less certain and
smaller in magnitude compared to the robust effects observed in models
of receptive vocabulary or in the baseline models that did not control
for Time 1 vocabulary. Comparison of the expressive vocabulary models
indicated that one should assign approximately equal weight to a model
with both lexical processing and Time 1 vocabulary and a model with only
Time 1 vocabulary. In contrast, the best predictors of Time 2 receptive
vocabulary were Time 1 receptive vocabulary, followed by lexical
processing and adult word count. Both processing and input provided
additional predictive information over and above Time 1 vocabulary, and
lexical processing had a larger effect than hourly adult word count.
Finally, we estimated the differences in the effects on receptive versus
expressive vocabulary, and the input effect was larger for receptive
vocabulary while the processing effect was *probably* larger for
receptive vocabulary.

The difference in results for expressive versus receptive vocabulary was
unexpected, given the reliable correlation between the two measures,
*r* = .81 at ages 2--5 [@PPVT4, p. 60]. Child who heard more words
from their caregivers could understand more words one year later, but
they could not necessarily produce more words. Why would language
exposure be more related to receptive than expressive vocabulary? The
differences for expressive and receptive vocabulary may simply reflect
differences in the tests used to measure them. Across the entire course
of the PPVT-4, the prompt remains the same (*show me X*). But over the
course of the EVT-2, the prompts change from *what is this* to include
prompts that demand metalinguistic knowledge (e.g., *tell me another
word for X*). Thus, it may be the case that the PPVT-4 measures only
receptive vocabulary, while the EVT-2 measures both expressive
vocabulary and metalinguistic ability. Lexical processing efficiency and
language input may be less related to metalinguistic ability than they
are to vocabulary size.

Alternatively, the different results for expressive and receptive
vocabulary may reflect the fact that recognition is easier than
production. Being able to name an object---to activate the word's
semantic representation and its phonological representation then carry
out a motor plan---demonstrates a greater sign of mastery than being
able to associate the word to an appropriate referent. The children who
heard more words had more experience and exposure to words, giving them
a broad base of shallow knowledge for word recognition. This
interpretation would suggest that measures of input diversity (input
types) would be even more predictive of future receptive vocabulary than
simple quantity. Support for this interpretation also comes from
@edwards2014dialect. Using structural equation modeling, they found a
direct relation between SES and receptive vocabulary, but only an
indirect relation between SES and expressive vocabulary; the relation
between SES and expressive vocabulary was mediated by receptive
vocabulary.

A similar line of reasoning applies to the processing effect. We
measured processing as response speed during a listening task, not a
naming task. It captures a child's ability to activate a word's semantic
representation in a timely manner. These demands are more clearly
related to our receptive vocabulary task, whereas the expressive
vocabulary task additionally required the child to talk. Nevertheless,
we concluded that faster processing *probably* predicted larger
expressive vocabularies. Approximately, 75% of the posterior samples
indicated a positive, non-null effect. Naming an object still requires
activation of a word's lexical representations, so it makes sense that
the lexical processing efficiency would still matter for expressive
vocabulary.

Throughout our analysis, we never observed a credible interaction effect
between lexical processing and language input. Word recognition
efficiency did not constrain the beneficial effects of language exposure
on future receptive vocabulary size. One interpretation of these
findings is that these children were *fast enough* at recognizing words
that processing did not impose more of a bottleneck on vocabulary
growth. Developmentally, that bottleneck may be observed in younger
children than those in this sample. The youngest children in this study
were 28 months, an age at which the average child produces about 500
words and recognizes at least 3 times that amount. In contrast, at 18
months the average child produces only about 50 words and recognizes
about 250 [@Wordbank]. After about 18 months, children’s vocabularies
start increasing rapidly. At this point, it may be the case that
processing efficiency no longer interacts with the quantity of language
input. More research on children from 18 to 30 months is needed to
evaluate this claim and to determine the time course of the relation
among processing efficiency, language input, and vocabulary growth.

Our study elaborated on the work of @Weisleder2013, but differed in
important ways. Notably, our sample included children who were older in
age, and we tested these children's vocabularies directly. Additionally,
we used LENA’s automated measures of adult word count, whereas that
study used LENA word counts from just the segments of recordings that
listeners had classified as child-directed. Limitations of our study
include its observational design and the relatively homogeneous
demographics of the families. This study was observational, so the
analyses here describe statistical relationships. We did not manipulate
language input, so we cannot establish causal links between language
input and other measures. Moreover, most
(`r medu_counts$high`/`r medu_counts$total`) children came from high
maternal-education families (i.e., college or graduate degrees), whereas
@Weisleder2013 recruited 29 children from low-SES families. The combined
SES and age differences make it difficult to compare these two studies
directly. Although most of the participants in our study were
demographically homogeneous, they varied in language input, vocabulary
size, and processing efficiency, so they provided an informative test of
how processing and input predict word learning. We found input and
processing had positive effects on receptive vocabulary growth, but had
little influence on expressive vocabulary growth.

Once we controlled for Time 1 vocabulary size, the effects of lexical
processing and language input became less robust and less certain,
especially for expressive vocabulary. It is essential that studies about
changes in vocabulary size obtain a baseline vocabulary measurement. Our
results were different when we measured predictors of vocabulary size as
compared to predictors of vocabulary growth. This is not surprising,
given that individual differences in vocabulary size are observed at 12
months or even earlier and increase with age. Unlike previous studies,
this study measured both receptive and expressive vocabulary---if only
one of these measures had been used as the sole measure of vocabulary
knowledge, we would have drawn different conclusions.

Our findings have important implications for interventions aimed at
increasing children’s vocabulary. First and foremost, these
interventions must start early. At Time 1, children in this study were
only 28 to 39 months of age. Still, for both receptive and expressive
vocabulary, vocabulary size at this young age was by far the strongest
predictor of vocabulary size one year later. Based on our results,
increasing the quantity of linguistic input for 3-year-olds is not going
to be an effective intervention strategy. Other research has suggested
that children attend to different features of their ambient language as
their language abilities develop [@Rowe2012_Long; cf. review in
@Lidz2015]. For children of this age, sheer quantity of language input
may not be as relevant or predictive as complexity or other features of
the child-directed speech. The lack of evidence for quantity effects
does not imply that quality is any more important or predictive of
vocabulary growth at this age, but it suggests that measures like adult
word count provide only a first approximation about the number of
informative examples and learning opportunities available to the child.
Language stimulation benchmarks and goals for children of this age, we
would conclude, are better framed as time spent on activities, such as
shared-book reading, or discussing events outside of the here and now
[@Rowe2012_Long], rather than focusing simply on increasing language
input.

These findings add to our understanding for vocabulary development. They
suggest that the early relations among processing efficiency, input, and
vocabulary size may decrease as all children become more efficient
language processors in their third year of life. While differences in
processing efficiency continue to exist, they may not create the
bottleneck on vocabulary growth that is observed for children from 18
to 24 months. They also add to the very large literature supporting
early intervention---children say their first words at about 12 to 14
months of age, but differences in vocabulary size even as early as 28
months are highly predictive of vocabulary growth one year later and are
far more important than language input or processing efficiency.
