---
csl: assets/apa.csl
bibliography: assets/refs.bib
output: github_document
...

```{r appendix-setup, include = FALSE}
library(knitr)
wd <- rprojroot::find_rstudio_root_file()
opts_knit$set(root.dir = wd)
opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r appendix-packages}
library(dplyr, warn.conflicts = FALSE)
library(rstanarm)
library(printy)
source("./R/utils.R")
```

```{r appendix-gca-tables}
load("./models/00_gca.RData")

# Clean up the fixed effect parameter names
fix_param_names <- . %>%
  stringr::str_replace("ot(\\d)", "Time^\\1^")  %>%
  stringr::str_replace(":", " &times; ") %>%
  stringr::str_replace(".Intercept.", "Constant") %>%
  stringr::str_replace("Subj", "Child")

# Assemble fixed effects table
df_fixef <- gca$model %>% 
  broom::tidy(effects = "fixed") %>% 
  rename(
    Parameter = term, Estimate = estimate, SE = std.error, 
    z = statistic, p = p.value) %>% 
  mutate(
    p = fmt_p_value(p), 
    Parameter = fix_param_names(Parameter)) %>% 
  mutate_at(vars(Estimate:z), funs(format_fixef_num)) %>% 
  rename(`_z_` = z, `_p_` = p, ` ` = Parameter)

# Create random effects table
df_ranef <- gca$model %>% 
  tidy_ranef_summary() %>% 
  # Format the numbers
  mutate_at(vars(vcov, sdcor), funs(format_fixef_num)) %>%
  mutate_at(vars( -var1, -grp, -vcov, -sdcor), funs(format_cor)) %>%
  sort_ranef_grps() %>%
  # Format variable names and group names
  mutate(
    var1 = var1 %>% fix_param_names() %>% fmt_replace_na("&nbsp;"),
    grp = grp %>% replace_same_as_last("&nbsp;") %>% fix_param_names()) %>% 
  rename(Group = grp, Parameter = var1, Variance = vcov, SD = sdcor)

# Correlation columns need names with characters so that pandoc can align them
names(df_ranef)[5:8] <- c("Correlations", "&nbsp;", " &nbsp;", "  &nbsp;")
```

```{r appendix-vocab-tables}
evt_t2_models <- readRDS("./models/03a_evt_gsv_with_t1.rds")
ppvt_t2_models <- readRDS("./models/03b_ppvt_gsv_with_t1.rds")

exp_model <- evt_t2_models$models$t1_slope
rec_model <- ppvt_t2_models$models$t1_input_slope

# How much to round each of the columns
round_by_2 <- c("mean", "sd", "2.5%", "25%", "50%", "75%", "97.5%", "Rhat")

create_model_table <- . %>% 
  summary() %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("Parameter") %>% 
  filter(Parameter != "mean_PPD", Parameter != "log-posterior") %>% 
  select(-mcse) %>% 
  purrr:::map_at(.at = round_by_2, .f = ~ fmt_fix_digits(.x, 2)) %>% 
  as_data_frame() %>% 
  mutate_at(vars(-Parameter), funs(fmt_minus_sign)) %>% 
  rename(`_N_~eff~` = n_eff, `_R_~hat~` = Rhat, Mean = mean, SD = sd)
```

Appendix
------------------------------------------------------------------------

Data and R scripts to reproduce this manuscript are available online
`https://github.com/tjmahr/mahr-edwards-2017`. (Exact numbers will not
reproduce because variability in Monte Carlo sampling.)

### Eyetracking Growth Curve Model

The tables below summarize the parameter point estimates for the
mixed-effects logistic regression model of eyetracking data. These
models were fit using maximum likelihood. The lme4 syntax used to fit
the model was 
`cbind(Target, Others) ~ t1 + t2 + t3 + (t1 + t2 + t3 | Child)`. Here
`t1`, `t2` and `t3` are orthogonal polynomials produced by `poly()` but
rescaled so that `t1` ranges from −.5 to .5.

#### Fixed effects

```{r appendix-gca-fixef, results = "asis"}
df_fixef %>% 
  knitr::kable(align = str_tokenize("lrrrr"))
```

#### Random effects

```{r appendix-gca-ranef, results = "asis"}
df_ranef %>% 
  knitr::kable(align = str_tokenize("llrrrrrr"))
```

### Vocabulary Models

The tables below summarize the posterior distribution of parameters from
the best-performing models. We obtained 4,000 samples from the posterior
distribution in each model. Recall that all variables were scaled to
have mean of 0 and an SD of 1. Two diagnostic measures are included.
_N_~eff~ estimates the effective number of independent samples from
the posterior. _R_~hat~ estimates the convergence of the MCMC sampling
chains; its values should be close to 1.00.

#### Expressive vocabulary

Likelihood: 

ExpTime2~_i_~ &#x223C; _Normal_(_µ_~_i_~, _σ_)  \
_µ_~_i_~ = _α_ + _β_~1~ExpTime1~_i_~ + _β_~2~Processing~_i_~

Priors: 

_α_ &#x223C; _Normal_(0, 5)  \
_β_~_j_~ &#x223C; _Normal_(0, 1)  \
_σ_ &#x223C; _HalfCauchy_(0, 5)

```{r appendix-exp-vocab, results = "asis"}
exp_model %>% 
  create_model_table() %>% 
  mutate(Parameter = create_math_predictors(Parameter, "expressive")) %>% 
  rename(` ` = Parameter) %>% 
  knitr::kable(align = str_tokenize("lrrrrrrrrr"))
```

#### Receptive vocabulary

Likelihood: 

RecTime2~_i_~ &#x223C; _Normal_(_µ_~_i_~, _σ_)  \
_µ_~_i_~ = _α_ + _β_~1~RecTime1~_i_~ + _β_~2~Input~_i_~ + _β_~3~Processing~_i_~  \

Priors: 

_α_ &#x223C; _Normal_(0, 5)  \
_β_~_j_~ &#x223C; _Normal_(0, 1)  \
_σ_ &#x223C; _HalfCauchy_(0, 5)

```{r appendix-rec-vocab, results = "asis"}
rec_model %>% 
  create_model_table() %>% 
  mutate(Parameter = create_math_predictors(Parameter, "receptive")) %>% 
  rename(` ` = Parameter) %>% 
  knitr::kable(digits = 2, align = str_tokenize("lrrrrrrrrr"))
```

### Unreported analyses

In response to peer review feedback, we performed two additional
versions of the main analyses that controlled for age. One used
vocabulary standard scores and another residualized age from the
vocabulary, processing and input measurements. Our inferences remained
the same in these analyses.

The original submission of this article reported a mediation analysis;
however, we removed because it did not address our research questions
and because there was a negligible effect of language input on
processing efficiency.
